---
title: "Dirty Data Project - Task 4 Halloween Candy Data"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: '2'
    highlight: tango
    df_print: paged
---

## Introduction
The boing boing candy dataset contains the results of a Halloween candy survey. The data covers three years (2015, 2016 and 2017) and includes information on the person completing the survey, such as age, gender and country. The ratings used were Joy, Despair, Meh or NA for the 2016 and 2017 data; Meh was not an option in 2015. The 2015 data also lacked information on the gender and country of the person completing the survey.
The original csv file required extensive cleaning and the main steps conducted are summarised below in the cleaning section. 


## Assumptions
A lot of columns were removed, these were columns which weren't required to answer the analysis questions. 

1. The timestamp column would not be required and was converted to a year column. 
2. The columns containing data on degrees of separation would not be required and were therefore removed. 
4. Other columns which didn't have the joy, despair rating used were removed e.g. "What is your favourite font?" and "Which day of the week do you prefer?". 
3. All the analysis questions are about candy so I thought about removing columns for items that aren't candy (DVDs, pharmaceuticals, glow sticks) but decided it would be more interesting to keep them in. 
4. The state_province_county column was removed from 2016 and 2017 data as its not required to answer the analysis questions and 
5. Internal id removed from 2017 data. 


## Cleaning

For full details see cleaning script comments. A summary of the main cleaning steps is provided here:

- tidied up column names (using janitor package and manually)
- ensured the non-candy columns (age, gender etc) were the same across the 3 original datasets to allow joining
- joined the three datasets into one using full_join
- combined columns which belonged together but were given different names in different years
- converted age column to integer and set a realistic age range
- used case_when and str_detect to fix country column 
- admired my beautiful clean dataset



## Load libraries and import data
```{r}
library(tidyverse)
library(here)

candy <- read_csv(here("clean_data/clean_candy_data.csv"))
```



## Question 1

**What is the total number of candy ratings given across the three years?**

```{r}
candy %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~sum(!is.na(.x)))) %>% 
  rowSums()

# Code to check answer
# sum(is.na(candy[6:123]))
```
There were 772,352 candy ratings given across the three years. 

To check this number the number of NAs was determined and subtracted from the total number of entries. 
118 candy columns *  9349 rows = 1,103,182 entries
Missing values (NAs) = 330830
1,103,183 - 330,830 = 772,352 which is our original answer!


## Question 2

**What was the average age of people who are going out trick or treating?**

```{r}
candy %>% 
  filter(going_out == "Yes") %>% 
  summarise(avg_age_trick_or_treaters = round(mean(age, na.rm = TRUE)))
```
The average age of trick or treaters is 35 which seems rather high. Perhaps this is parents who are going out with their children. 

## Question 3

**What was the average age of people who are not going out trick or treating?**

```{r}
candy %>% 
  filter(going_out == "No") %>% 
  summarise(avg_age_trick_or_treaters = round(mean(age, na.rm = TRUE)))
```
The average age of people not going out trick or treating was 39. I can't think of anything interesting to say about this answer.


## Question 4

**For each of joy, despair and meh, which candy bar received the most of these ratings**

Assumption -  this means which candy received highest number of joy ratings, the highest number of meh ratings etc. 

```{r}
candy %>% 
  pivot_longer(cols = 6:123, names_to = "candy", values_to = "rating") %>% 
  group_by(candy) %>% 
  filter(rating == "JOY") %>% 
  summarise(total_ratings = n()) %>% 
  filter(total_ratings == max(total_ratings))

candy %>% 
  pivot_longer(cols = 6:123, names_to = "candy", values_to = "rating") %>% 
  group_by(candy) %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(total_ratings = n()) %>% 
  filter(total_ratings == max(total_ratings))


candy %>% 
  pivot_longer(cols = 6:123, names_to = "candy", values_to = "rating") %>% 
  group_by(candy) %>% 
  filter(rating == "MEH") %>% 
  summarise(total_ratings = n()) %>% 
  filter(total_ratings == max(total_ratings))


```
I wasted an extremely long time on this question by trying to answer it without pivoting :(
Any full sized candy bar brings the most joy, while broken glow stick bring the most despair which isn't that surprising. Lollipops get the highest meh rating which was a little surprising, I would have thought they would have been more popular. 






```{r}

highest_num_of_ratings <- function(dataset, JOY_DESPAIR_MEH_NA){
 
    dataset %>% 
    pivot_longer(cols = 6:123, names_to = "candy", values_to = "rating") %>% 
    group_by(candy) %>% 
    filter(rating == JOY_DESPAIR_MEH) %>% 
    summarise(total_ratings = n()) %>% 
    filter(total_ratings == max(total_ratings))

}


highest_num_of_ratings(candy, "JOY")
highest_num_of_ratings(candy, "DESPAIR")
highest_num_of_ratings(candy, "MEH")


gender_results <- as.data.frame(t(gender_results)) %>% 
  rownames_to_column() %>% 
  cbind(sex) %>% 
  rename(score = V1, sweet = rowname, gender = sex) %>% 
  relocate(gender, .before = sweet) %>% 
  filter(score == max(score))
  
  gender_results
}

# Creating empty dataframe 
most_popular_candy_by_gender <- data.frame(
  gender = as.character(),
  sweets = as.character(),
  score = as.numeric()
  ) %>% 
# Adding results from running the function with each gender to dataframe
    rbind(high_score_by_gender(candy, "female"), high_score_by_gender(candy, "male"), 
        high_score_by_gender(candy, "i'd rather not say"), high_score_by_gender(candy, "other"))

```


## Question 5

**How many people rated Starburst as despair?**

```{r}
candy %>% 
  filter(starburst == "DESPAIR") %>% 
  summarise(starburst_as_despair = n())
```
Personally I am not a fan of starburst (or Opal Fruits as they were when I was a kid) so I feel this number should be higher!


## Question 6

**What was the most popular candy bar by the -1, 0, 1 rating system for each gender in the dataset?**

```{r}
# These JOY, DESPAIR, MEH strings won't be identified in the non-candy columns (e.g. if "meh" was part of a country name) as they are capitalised and the other columns are lower or sentence case. 

candy <- candy %>% 
  mutate_all(~str_replace(., "JOY", "1")) %>% 
  mutate_all(~str_replace(., "DESPAIR", "-1")) %>% 
  mutate_all(~str_replace(., "MEH", "0")) %>% 
  mutate(across((6:123), as.numeric))

# remove this once cleaning script has been rerun and analysis data reloaded
candy$gender <- str_to_lower(candy$gender)


# The function below filters by gender, then sums each of the columns
# The results are converted into a dataframe, transposed and filtered to reveal the top scoring candy. 

high_score_by_gender <- function(dataset, sex){
# sex options are: female, male, i'd rather not say, other
  gender_results <- dataset %>% 
  filter(gender == sex) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
 
  gender_results <- as.data.frame(t(gender_results)) %>% 
  rownames_to_column() %>% 
  cbind(sex) %>% 
  rename(score = V1, sweet = rowname, gender = sex) %>% 
  relocate(gender, .before = sweet) %>% 
  filter(score == max(score))
  
  gender_results
}

# Creating empty dataframe 
most_popular_candy_by_gender <- data.frame(
  gender = as.character(),
  sweets = as.character(),
  score = as.numeric()
  ) %>% 
# Adding results from running the function with each gender to dataframe
    rbind(high_score_by_gender(candy, "female"), high_score_by_gender(candy, "male"), 
        high_score_by_gender(candy, "i'd rather not say"), high_score_by_gender(candy, "other"))

most_popular_candy_by_gender
           
```
I realise now I could have very easily answered this question by using pivot_longer but I love my function and it took me ages so I've left it in. It also displays all the results together in a nice dataframe. 

As for the results, nothing interesting here, full sized candy bars popular across all genders. 

## Question 7

**What was the most popular candy bar in each year?**

```{r}
high_score_by_year <- function(dataset, yr){
# year options are: 2015, 2016 and 2017
  year_results <- dataset %>% 
  filter(year == yr) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
# Tried to keep the pipe going but couldn't get the as.data.frame(t()) function to work so left the above and below code separate 
  year_results <- as.data.frame(t(year_results)) %>% 
  rownames_to_column() %>% 
  cbind(yr) %>% 
  rename(score = V1, sweet = rowname, year = yr) %>% 
  relocate(year, .before = sweet) %>% 
  filter(score == max(score))
  
  year_results
}

# Creating empty dataframe 
most_popular_candy_by_year <- data.frame(
  year = as.numeric(),
  sweets = as.character(),
  score = as.numeric()
  ) %>% 
# Adding results from running the function with each year to dataframe
    rbind(high_score_by_year(candy, 2015), high_score_by_year(candy, 2016), 
        high_score_by_year(candy, 2017))

most_popular_candy_by_year
```

I adapted my beautiful function to work with year, again pivot_longer would have provided a much quicker and easier answer. 
Full sized candy bars proving popular across the years. 


## Question 8

**What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries**

```{r}
high_score_by_country <- function(dataset, cntry){
  country_results <- dataset %>% 
  filter(country == cntry) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
# Tried to keep the pipe going but couldn't get the as.data.frame(t()) function to work so left the above and below code separate 
  country_results <- as.data.frame(t(country_results)) %>% 
  rownames_to_column() %>% 
  cbind(cntry) %>% 
  rename(score = V1, sweet = rowname, country = cntry) %>% 
  relocate(country, .before = sweet) %>% 
  filter(score == max(score))
  
  country_results
}

# Creating empty dataframe 
most_popular_candy_by_country <- data.frame(
  country = as.character(),
  sweets = as.character(),
  score = as.numeric()
  ) 

# Adding results from running the function with countries to dataframe
most_popular_candy_by_country <- most_popular_candy_by_country %>% 
rbind(high_score_by_country(candy, "usa"), high_score_by_country(candy, "canada"), 
        high_score_by_country(candy, "uk"))

most_popular_candy_by_country



```

In the UK cash rules, here cash is rated higher than the full sized candy bar. 
I spend a very, very long time try to get the function to take in a vector of countries as an argument. I added an if else statement at the cbind() line which worked on length of the cntry argument but ran into errors. I understood why I was getting the error but didn't manage to fix it. The code of "all other countries" is below. 



```{r}
# Creating a dataframe of countries other than US, UK and Canada
all_other_countries_df <- candy %>% 
  filter(!country %in% c("usa", "canada", "uk")) %>% 
  filter(!is.na(country))
# Extracting countries into a vector
all_other_countries_vec <- all_other_countries_df$country


high_score_by_multiple_countries <- function(dataset, cntry){
  country_results <- dataset %>% 
  filter(country == cntry) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
# Tried to keep the pipe going but couldn't get the as.data.frame(t()) function to work so left the above and below code separate 
  country_results <- as.data.frame(t(country_results)) %>% 
  rownames_to_column() %>% 
  rename(score = V1, sweet = rowname) %>% 
  filter(score == max(score))
  
  country_results
}

high_score_by_multiple_countries(candy, all_other_countries_vec)






```







```{r}
high_score_by_NA_country <- function(dataset){
  country_results <- dataset %>% 
  filter(is.na(country)) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
# Tried to keep the pipe going but couldn't get the as.data.frame(t()) function to work so left the above and below code separate 
  country_results <- as.data.frame(t(country_results)) %>% 
  rownames_to_column() %>% 
  rename(score = V1, sweet = rowname) %>% 
  filter(score == max(score))
  
  country_results
}

high_score_by_NA_country(candy)
```








```{r}
top_3_by_country <- function(dataset, cntry){
  top3_country <- dataset %>% 
  filter(country == cntry) %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~ sum(., na.rm = TRUE)
    )
  )
# Tried to keep the pipe going but couldn't get the as.data.frame(t()) function to work so left the above and below code separate 
  top3_country <- as.data.frame(t(top3_country)) %>% 
  rownames_to_column() %>% 
  cbind(cntry) %>% 
  rename(score = V1, sweet = rowname, country = cntry) %>% 
  relocate(country, .before = sweet) %>% 
  arrange(desc(score)) %>% 
  head(3)
    
    #filter(score == slice_max(score, 3))
  
  top3_country
}

# Creating empty dataframe 
top3_candies_by_country <- data.frame(
  country = as.character(),
  sweets = as.character(),
  score = as.numeric()
  ) 

# Adding results from running the function with countries to dataframe
top3_candies_by_country <- top3_candies_by_country %>% 
rbind(top_3_by_country(candy, "usa"), top_3_by_country(candy, "canada"), 
        top_3_by_country(candy, "uk"))

top3_candies_by_country


top_3_by_country(candy, "uk")

# Would be nice to add a rank but I'm done
```

