---
title: "Dirty Data Project - Task 4 Halloween Candy Data"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: '2'
    highlight: tango
    df_print: paged
---

## Introduction



## Assumptions



## Cleaning


## Load libraries and import data
```{r}
library(tidyverse)
library(here)

candy <- read_csv(here("clean_data/clean_candy_data.csv"))
```



## Question 1

**What is the total number of candy ratings given across the three years?**

```{r}
candy %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~sum(!is.na(.x)))) %>% 
  rowSums()


```



## Question 2

**What was the average age of people who are going out trick or treating?**

```{r}
candy %>% 
  filter(going_out == "Yes") %>% 
  summarise(avg_age_trick_or_treaters = round(mean(age, na.rm = TRUE)))
```

## Question 3

**What was the average age of people who are not going out trick or treating?**

```{r}
candy %>% 
  filter(going_out == "No") %>% 
  summarise(avg_age_trick_or_treaters = round(mean(age, na.rm = TRUE)))
```

## Question 4

**For each of joy, despair and meh, which candy bar received the most of these ratings**

Assuming this means - which candy received highest number of joy ratings etc. 

```{r}
candy %>% 
  summarise(across(
    .cols = 6:123,
    .fns = ~sum(.x == "JOY")))



candy %>% 
  select(abstained_from_m_and_m_ing) %>% 
  group_by(abstained_from_m_and_m_ing) %>% 
  summarise(count = n())


candy %>% 
  group_by(abstained_from_m_and_m_ing)
  summarise(across(
    .cols = 6:123,
    .fns = ~ n()
  )
  )


  
  
  

```

## Question 5

**How many people rated Starburst as despair?**

```{r}
candy %>% 
  group_by(starburst) %>% 
  summarise(count = n())
```

