---
title: "Dirty Data Task 1 Decathlon"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: '2'
    highlight: tango
    df_print: paged
---

why did i chose this task
intro to the data

```{r}
library(tidyverse)
library(here)
```

```{r}
here::here()

decathlon <- read_csv(here("clean_data/clean_decathlon_data.csv"))
```



1. Who had the longest long jump seen in the data?
```{r}
decathlon %>% 
  filter(long_jump == max(long_jump))
```
Bryan Clay had the longest result for the long jump from the data. This jump was 
at the 2004 Olympic Games and was a personal best at the time. Clay went on to win
the silver medal in this competition. 

2. What was the average 100m time in each competition?

```{r}
decathlon %>% 
  group_by(competition) %>% 
  summarise(avg_100m_time = round(mean(x100m), 2))
```
The Olympic Games has a faster average time for the 100m than Decastar. I don't 
think this is a surprising result. The Olympics Games is a higher profile 
competition and athletes design four year training programs around the Olympic 
schedule to ensure they peak for this competition. 

3. Who had the highest total points across both competitions?

Assumption - highest total points across both competitions means the points are 
summed for each athlete (that competed in both competitions) and the athelte with
the highest points identified. 

```{r}
decathlon %>% 
  group_by(athlete) %>% 
  select(athlete, points) %>% 
  mutate(total_points = sum(points)) %>% 
  filter(total_points == max(total_points))
```

