---
title: "Dirty Data Project Task 1 Decathlon"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: '2'
    highlight: tango
    df_print: paged
---

why did i chose this task intro to the data

```{r}
library(tidyverse)
library(here)
```

```{r}
here::here()

decathlon <- read_csv(here("clean_data/clean_decathlon_data.csv"))
```

1.  Who had the longest long jump seen in the data?

```{r}
decathlon %>% 
  filter(long_jump == max(long_jump))
```

Bryan Clay from the US had the longest result for the long jump from the
data. This jump was at the 2004 Olympic Games and was a personal best at
the time. Clay went on to win the silver medal in this competition.

2.  What was the average 100m time in each competition?

```{r}
decathlon %>% 
  group_by(competition) %>% 
  summarise(avg_100m_time = round(mean(x100m), 2))
```

The Olympic Games has a faster average time for the 100m than Decastar.
I don't think this is a surprising result. The Olympics Games is a
higher profile competition and athletes design four year training
programs around the Olympic schedule to ensure they peak for this
competition.

3.  Who had the highest total points across both competitions?

Assumption - highest total points across both competitions means the
points are summed for each athlete (that competed in both competitions)
and the athlete with the highest points identified.

```{r}
decathlon %>% 
  group_by(athlete) %>% 
  summarise(total_points = sum(points)) %>% 
  filter(total_points == max(total_points))
```

Roman Sebrle, an athlete from the Czech Republic, has the highest total
points across both competitions. He won the 2004 Olympics and the 2004
Decastar event which is where this data comes from.

4.  What was the shot-put scores for the top three competitors in each
    competition?

Assumption - the question means the shot-put distance since it is not
possible to calculate the "shot-put scores" from this data.

```{r}
decathlon %>% 
# Slicing the 6 lowest numbers in rank column to catch the 1, 2, 3 ranked positions in each competition  
  slice_min(rank, n = 6) %>% 
  select(athlete, shot_put, rank, competition) %>% 
  arrange(competition, rank)

```

The first, second and third places in both competitions were taken by
the same men in the same order. The shot-put distances in the Olympics
are all higher than those thrown in the Decastar competition.

5.  What was the average points for competitors who ran the 400m in less
    than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
# A quick check showed no one ran exactly 50 seconds so all athletes will be caught by >50 and <50

avg_points_slower_athletes<- decathlon %>% 
  filter(x400m > 50) %>% 
  summarise(avg_points_slower_athletes = round(mean(points), 0)) %>% 
  pull()

avg_points_faster_athletes <- decathlon %>% 
  filter(x400m < 50) %>% 
  summarise(avg_points_faster_athletes = round(mean(points), 0)) %>% 
  pull()

tibble(
  "x400m_speed" = c("less than 50 sec", "more than 50 sec"),
  "average_points" = c(avg_points_slower_athletes, avg_points_faster_athletes)
)
```

The athletes who are slower at the 400 m have a lower points average
than those who are faster.
